### 2014-01-07 paper/cover.tex
added initial paper draft and initial port to PySide

\title{Detecting Photo Manipulation using Reflections on Curved Surfaces}
\author{Laura A. Moss}
\date{}
\maketitle

\newpage

\subsection*{Abstract}
Abstract goes here