### 2014-01-07 paper/intro.tex

\section{Introduction}

\subsection{Image Forensics}
importance

Emerging field because of advances in processing power, which benefits both image manipulation and its detection

examples of curved reflective surfaces in the real world