### 2014-01-17 paper/intro.tex
more paper work

\section{Introduction}

\subsection{Image Forensics}
importance - journalism, scientific research

how long has the field been around?
Emerging field because of advances in processing power, which benefits both image manipulation and its detection
	examples of research

Most of the reflective surfaces that might be found in real photos are planar. Some common examples are bodies of still water, mirrors, and glass windows. Non-planar reflective surfaces are less common, but can be found in the form of store security mirrors, works of art, and even the human eye.