### 2014-02-20 paper/cover.tex
added details about camera parameters and counterforensics

\title{Image Forensics using Reflections on Curved Surfaces}
\author{Laura A. Moss}
\date{}
\maketitle

\newpage

\subsection*{Abstract}
Image forensics is the study of finding hidden information in photographs. This emerging field encompasses techniques for determining the source of an image, detecting forgery, and detecting steganography. Because pictures are used as hard evidence in journalism, scientific research, and criminal cases, having a strong arsenal of methods for detecting forgery is crucial. Many such methods rely on geometric properties of the image's content, such as mirrors. This research builds on existing methods for finding inconsistencies in reflections on planar mirrors by exploring techniques that would work for spherical mirrors.